---
title: "HW1"
author: "Gabriel Nespoli | Fernando Crema | Mauricio Fadel Argerich"
date: "October 9, 2016"
output: html_document
---

## Part I: R Syntax & Functions
### a.
The command creates a vector with 4 elements that are characters.
```{r}
vector1 <- c("5", "12", "7", "32")
```
The result of the command will be 7 because it's sorted alphabetically.
```{r}
max(vector1)
```
The vector is sorted alphabetically (not using the numeric value of each element), so the result is: 12, 32, 5, 7.
```{r}
sort(vector1)
```
The command produces an error because it's not possible to sum a vector which elements are characters.
```{r eval=FALSE}
sum(vector1)
```

### b.
The commands don't work. Because the first element of the vector is a character, all the elements in the vector are changed to the most "generic" type of data of its elements which in this case is characters. So when we try to add the second and third element, R tells us it's not possible to add non-numeric arguments, because they're characters.
```{r eval=FALSE}
vector2 <- c("5", 7, 12) 
vector2[2] + vector2[3]
```
The commands work and the result is 19 (7+12). A mainframe can contain different kind of data types, so even the first element is a character, the other elements are number and adding the second and third element works fine.
```{r}
dataframe3 <- data.frame(z1 = "5", z2 = 7, z3 = 12) 
dataframe3[1,2] + dataframe3[1,3]
```
Each element of a list is a list, when we do list4[2] we get the second element of the list (with its name and value), while doing list4[[2]] returns the value of the element. This is why the second command in the code chunk works while the last one doesn't.
```{r eval=FALSE}
list4 <- list(z1 = "6", z2 = 42, z3 = "49", z4 = 126) 
list4[[2]] + list4[[4]]
list4[2] + list4[4]
```

### c.
From 1 to 10000 in increments of 372.
```{r}
seq(1,10000,372)
```
Sequence of 50 numbers from 1 to 10000.
```{r}
seq(1,10000,length.out = 50)
```

### d.
This command repeats the whole vector 3 times in order.
```{r}
rep(1:3, times = 3)
```
While this command repeats each element of the vector 3 times (before passing on to the next element).
```{r}
rep(1:3, each = 3)
```

## Part V: A naïve version of the Naïve Bayes Classifier
The exercise gives us a table of the historical events of four weather parameters (outlook, temperature, humidity, wind) and, based in these weather conditions, we played tennis or not. Then, knowing the weather condition for today, it asks us if we will play tennis. In other words, based in our past experience with the given condition, i.e. Outlook = Sunny, Temperature = Cool, Humidity = High , Wind = Strong, what is most propable evidence expected?

We start to compute the total "Yes" and "No" probability. In overall table, how many times we played tennis and how many we did not:
```{r eval=FALSE}
P(Play = Yes) = 9/14
P(Play = No) = 5/14
```

Considering this results, we calculate how many times we played tennis and how many we didn't for each hypothesis. For example, the conditional probability of playing tennis in a sunny day is:

```{r eval=FALSE}
P(Outlook = Sunny | Play = Yes) = 2/9
P(Outlook = Sunny | Play = No) = 3/5
```

For the others hypothesis, we have:

Outlook
```{r eval=FALSE}
P(Outlook = Overcast | Play = Yes) = 4/9
P(Outlook = Overcast | Play = No) = 0/5
P(Outlook = Rain | Play = Yes) = 3/9
P(Outlook = Rain | Play = No) = 2/5
```

Temperature
```{r eval=FALSE}
P(Temperature = Hot | Play = Yes) = 2/9
P(Temperature = Hot | Play = No) = 2/5
P(Temperature = Mild | Play = Yes) = 4/9
P(Temperature = Mild | Play = No) = 2/5
P(Temperature = Cool | Play = Yes) = 3/9
P(Temperature = Cool | Play = No) = 1/5
```

Humidity
```{r eval=FALSE}
P(Humidity = High | Play = Yes) = 3/9
P(Humidity = High | Play = No) = 4/5
P(Humidity = Normal | Play = Yes) = 6/9
P(Humidity = Normal | Play = No) = 1/5
```

Wind
```{r eval=FALSE}
P(Wind = Strong | Play = Yes) = 3/9
P(Wind = Strong | Play = No) = 3/5
P(Wind = Weak | Play = Yes) = 6/9
P(Wind = Weak | Play = No) = 2/5
```

Considering these evidences hypothesis, to evaluate the probability of playing tennis or not, we have to calculate the expressions:

```{r eval=FALSE}
argmax{P(Yes | Outlook = Sunny, Temperature = Cool, Humidity = High , Wind = Strong),
P(No | Outlook = Sunny, Temperature = Cool, Humidity = High , Wind = Strong)}
```

Calculating each argument separetely we have:
```{r eval=FALSE}
P(Yes | (Sunny, Cool, High , Strong)) = P(Yes)*[P(Sunny|Yes)*P(Cool|Yes)*P(High|Yes)*P(Strong|Yes)]
```
And
```{r eval=FALSE}
P(No | (Sunny, Cool, High , Strong)) = P(No)*[P(Sunny|No)*P(Cool|No)*P(High|No)*P(Strong|No)]
```

Which gives us:
```{r eval=FALSE}
P(Yes | (Sunny, Cool, High , Strong)) = (9/14)*(2/9 * 3/9 * 3/9 * 3/9)
P(No | (Sunny, Cool, High , Strong)) = (5/14)*(3/5 * 1/5 * 4/5 * 3/5)
```

Naming P(Yes | (Sunny, Cool, High , Strong)) P_Yes and P(No | (Sunny, Cool, High , Strong)) P_No and calculating, we have:

```{r}
P_Yes <- (9/14)*(2/9 * 3/9 * 3/9 * 3/9)
P_Yes

P_No <- (5/14)*(3/5 * 1/5 * 4/5 * 3/5)
P_No
```

As P_No is greater than P_Yes, we will not play tennis tomorrow.